\tool: A Multi-Mode Attack Model of BadUSB

### 2021-01-22 WOOT/Files/related_work.tex
revision of defense related work

\subsection{USB Attack}
\label{subsec:usb_attack}
While many USB based attacks have been proposed, these attacks are all tied to the environment of the target devices. In addition, limited to the USB 2.0\cite{usb20} specification, there is no video transferring interface, the control of attack processing is impossibly handled by the attackers without additional processing signal received.\hongyi{What?} One way is to masquerade protocol of attack devices, e.g., Rubbery Ducky \cite{rubber},\cite{badusb}, \cite{rubberducky2020}, \cite{usbbypassing}, \cite{iseeyou}and USBderivery\cite{usbdriver}, when they are connected to the host, they will hide their own functionality with the re-written firmware, that works because of HID enumeration does not verify the firmware of devices. Software running on the host will exfiltrate data or receive commands via the RF interface in \cite{turnip} after modifying the USB table as an additional RF transmitter. What is more, the host's USB software stack generally expects devices to confirm the USB standard, \cite{facedancer} and \cite{syzkaller} take a number of vulnerabilities to the kernel-model with the help of fuzzing technique. These attacks are all using the vulnerabilities of USB protocol, but due to the limitations of USB 2.0, there are no port for video transferring in the devices with USB 2.0, so it is impossible to obtain the real-time screen information of the host, only environment fitted commands and scripts will work in these scenarios.

@@

\subsection{USB Defence Security}
\label{subsec:usb_defence}
There are already many defenses proposed against BadUSB attacks and a comprehensive investigation of previous work.\cite{sok}.
BadUSB attack requires `D-' and `D+' which have been defined by USB protocol since USB 1.0\cite{usb10}.
Without these pins, data can't be transferred via USB cable. USB Condom \cite{Condom} proposed a hardware-based solution to prevent data from being transmitted on data channel by adding special hardware between the USB device and host. This hardware can cut off the `D-' and `D+' and allow only power pins.
However, this method poses a great challenge to easy-to-use property, and once deployed, it will block all USB functions other than charging.

Under the premise of ensuring the normal operation of all USB pins, some work improve the security of connection establishment.
Windows Defender ATP\cite{windenfenderwhite} maintains a whitelist of USB devices, only if one USB device is in the whitelist can it interact with host through data channel. This prevents all interactions between untrusted devices and the host. But this requires users to have a certain safety awareness. For example, a naive user may add the USB device from unknown sources to his/her whitelist without precaution. Mohammadmoradi et al\cite{mohammadmoradi2018making} propose a strategy to generate a whitelist automatically. Based on the features and functions of the USB devices, Mohammadmoradi et al generate unique fingerprint to each device and classify them. There are also some work like TMSUI\cite{yang2015tmsui}, which is based on the unreliable information provided by USB devices during enumeration to defend against BadUSB attack. 
However, this work relies on the information provided by the device during enumeration. Some USB devices may hide their real functionality from the user.
GoodUSB\cite{tian2015defending} will report the functionality and identity of the device to user and user needs to explicitly authorize. On this basis, GoodUSB loads the corresponding driver for the device, and device can not request other drivers during enumeration or usage. This restricts the device to load only the drivers allowed by the user, but does not prevent the device from using available drivers to attack. For example, a malicious keyboard can type commands without the user's knowledge after enumeration. Because from the user's perspective, he/she must permit a USB-based keyboard the corresponding drivers by authorizing them him/herself. But once the drivers are loaded, the malicious USB-based keyboard can also automatically type in commands.

When the USB device is normally enumerated and can interact with the host, there are also some defenses. 
Neuner et al.\cite{neuner2018usblock} prevent USB key injection attacks from malicious flash drive by analyzing temporal characteristics of BadUSB-like attacks. This defense mechanism is effective as the attacker cannot obtain the screen of the host using BadUSB. In this case, the malicious device can only inject instructions in a very short time to reduce the risk of being discovered. It leads to the typing characteristics of BadUSB devices to be abnormal and detectable. Apart from key injection attacks, malicious file injection can also be detected. Pham et al. \cite{pham2010optimizing} and odix\cite{OLEA} propose scanning function for files on portable USB devices, and can prevent malicious code from attacking host. What's more, in GoodUSB, VM is deployed in host to interact with the USB device to detect malicious behavior like automatic execution of code. In addition to injection attacks, data theft attack is also one of the focuses of the academic community. When the screen data of the mobile device is stolen while charging, the power consumption of the mobile device will change. So Meng et al\cite{meng2018252} detect potential attacks by monitoring and analyzing changes in the usage of CPU and GPU.

Some composite anti-virus systems, like Metascan\cite{OPSWAT}, incorporates some defenses. Metascan can verify digital signature of USB device, sanitize the malicious files in the device to stop malware, and deploy kiosks to protect security-sensitive organizations. However, Metascan also warned users to be cautious of the device plugged into the host, since there is no way to ensure absolute security.

\linyou{logic connection and pros/cons}

>>>>

\subsection{USB Attack}
\label{subsec:usb_attack}
\chaozu{DoS(Fuzzing), Code injection, HID emulation, JFC, our work}
While many USB based attacks have been proposed, these attacks are all tied to the environment of the target devices. In addition, limited to the USB 2.0\cite{usb20} specification, there is no video transferring interface, the control of attack processing is impossibly handled by the attackers without additional processing signal received.\hongyi{What?} One way is to masquerade protocol of attack devices, e.g., Rubbery Ducky \cite{rubber},\cite{badusb}, \cite{rubberducky2020}, \cite{usbbypassing}, \cite{iseeyou}and USBderivery\cite{usbdriver}, when they are connected to the host, they will hide their own functionality with the re-written firmware, that works because of HID enumeration does not verify the firmware of devices. Software running on the host will exfiltrate data or receive commands via the RF interface in \cite{turnip} after modifying the USB table as an additional RF transmitter. What is more, the host's USB software stack generally expects devices to confirm the USB standard, \cite{facedancer} and \cite{syzkaller} take a number of vulnerabilities to the kernel-model with the help of fuzzing technique. These attacks are all using the vulnerabilities of USB protocol, but due to the limitations of USB 2.0, there are no port for video transferring in the devices with USB 2.0, so it is impossible to obtain the real-time screen information of the host, only environment fitted commands and scripts will work in these scenarios.

@@

\subsection{USB Defence Security}
\label{subsec:usb_defence}
There are already many defenses proposed against BadUSB attacks and a comprehensive investigation of previous work.\cite{sok}.

From the hardware perspective, BadUSB attack requires `D-' and `D+' pins which are defined by protocol to transmitting data.
Without these pins, data can't be transferred via USB cable. Based on this fact, USB Condom \cite{Condom} is a hardware solution to block data channel by adding blocker in the connector. This blocker can cut off the `D-' and `D+' connection while leave power pins intact.
However, this method poses a great challenge to plug-and-play property of USB, as once it is deployed, it will stop all USB functions other than charging.

Under the premise of ensuring the full functionality of USB devices, there have been some works trying to improve the security during connection establishment.
Windows Defender ATP\cite{windenfenderwhite} maintains a whitelist of USB devices, only devices on the whitelist are allowed to communicate with the host. This prevents all potential attacks from untrusted devices. But this requires users to have a certain safety awareness and technical background to maintain a valid whitelist. For example, a naive user may add the USB device from unknown sources to his/her whitelist without precaution. In order to overcome this drawback, Mohammadmoradi et al\cite{mohammadmoradi2018making} proposes a strategy to generate such a whitelist automatically. This strategy first generates a unique fingerprint for each device based on its functionality. Then these fingerprints are used to maintain a safe and valid whitelist of USB devices. Apart from whitelist approach, there are also works like TMSUI\cite{yang2015tmsui}, which identifies the unreliable information provided by USB devices and thus defend against BadUSB attack. However, this work relies on the information provided by the device during enumeration. Some modified USB devices may hide their real functionality from the host.

To solve this flaw, GoodUSB\cite{tian2015defending} will report the functionality claimed by the device to user and let user decides whether to authorize. When a device is plugged in, GoodUSB will load its driver but limit its functionality until a series of authorizations is completed. These authorizations are designed to be performed manually. As BadUSB is normally unable to obtain the video stream of the host, it is impossible for an attacker to complete these authorization with automatic script. Thus this defense is sufficient for normal BadUSB attack. But if the attacker has the access to victim's screen, GoodUSB will be bypassed easily as the attacker can just complete these authorization manually and perform subsequent attacks.

After the USB enumeration and driver loading, there are also works trying to archive `defense in depth' against BadUSB attack.
Neuner et al.\cite{neuner2018usblock} prevents BadUSB attacks from malicious flash drive by analyzing the temporal characteristics of BadUSB-like attacks. This defense mechanism is effective because the attacker cannot obtain the screen of the host using just BadUSB. In this case, the malicious device can only inject key stoke in a very short time to reduce the risk of being discovered. This defect causes the typing characteristics of BadUSB to be detectable.  Pham et al. \cite{pham2010optimizing} and odix\cite{OLEA} propose scanning function for files on portable USB devices,\hongyi{What?} and can prevent malicious code from attacking host. What's more, in GoodUSB, a VM is deployed in host to as a honeypot to detect and stop malicious behavior of USB devices.

In addition to injection attacks, data theft attack is also one of the focuses of academic community. As mentioned in Section~\ref{subsec:usb_attack}, there exist an attack called JFC (juicy film charging)\cite{JFC} which abuses the MHL standard to steal video stream from the victim. In order to mitigate this issue, Meng et al proposed a statistical model using status like GPU/CPU usage to detect JFC attack\cite{meng2018252}.

To summarize, there exists a clear trade-off between the effectiveness and the plug-and-play property. Though hardware disabling solution like USB Condom archives the almost absolute security, the functionality of USB is scarified. Other solutions like GoodUSB or whitelist are either bypassable or insufficient under certain cases. There are many challenges to overcome before an optimal defense is formulated.\hongyi{Are there better way to express this?}
